Making Sequential Programs Deterministic

### 2019-04-01 experiments.tex
Fix capitalization of NonDex

\todo{check if the bugs discovered by \theDeterminismChecker have test coverage}
\todo{We investigated the tests covering code that \theDeterminismChecker reported as nondeterministic.
If code is covered by tests, NONDEX couldn't discover it. Else, its a limitation of NONDEX}.
This indicates that the additional programmer effort required by \TheDeterminismChecker is justified
as our tool discovers non-trivial issues and doesn't depend on code coverage or other runtime heuristics.

@@

All the flakiness that NONDEX found was due to 7 methods in 3 classes:
\<Class.getDeclaredFields>, \<Class.getDeclaredMethods>, \<Class.getFields>, \<DateFormatSymbols.getZoneStrings>,
\<HashMap.entrySet>, \<HashMap.keySet>, and \<HashMap.values>.  
We provide samples of the annotated source code form these benchmarks and the errors as reported by \TheDeterminismChecker 
in Figures~\ref{code-reflectasm} and~\ref{code-actiongenerator}.
\begin{figure}
    \begin{verbatim}
    static public FieldAccess get (Class type) {
        @Det ArrayList<@Det Field> fields = new ArrayList<Field>();
        ...
        while (nextClass != Object.class) {
            @Det Field @Det[] declaredFields = nextClass.getDeclaredFields();
            ...
        }
    }
    [ERROR] FieldAccess.java:[102,87] [assignment.type.incompatible] 
    incompatible types in assignment.
    found   : @Det Field @OrderNonDet []
    required: @Det Field @Det []
    \end{verbatim}
    \caption{Annotated source code from reflectasm.}
    \label{code-reflectasm}
\end{figure}
\begin{figure}
    \begin{verbatim}
    protected SimpleDataEvent createNextEvent() {
      for (@Det Entry<@Det String, @Det FieldType> entry : fields.entrySet()) {
          ...
      }
      ...
    }
    [ERROR] DataDictionaryEventSource.java:[73,78] [enhancedfor.type.incompatible] 
    incompatible types in enhanced for loop.
    found   : @NonDet Entry<@Det String, @Det FieldType>
    required: @Det Entry<@Det String, @Det FieldType>
    \end{verbatim}
    \caption{Annotated source code from ActionGenerator.}
    \label{code-actiongenerator}
\end{figure}
As can be seen from these figures, annotating local arrays and collections
has been sufficient to discover the sources of nondeterminism in most of the projects.
The rest of the projects required annotating the return types of methods.

For the benchmark \<commons-lang>, \cite{nondex} report multiple failing tests (14 flaky tests) due to calls
to \<Class.getDeclaredFields>. Upon inspecting the source code, we find that all of invocations to
\<Class.getDeclaredFields> are made from 5 locations. This means that annotating 5 lines of source code
would have been sufficient to identify the nondeterminism. Having said that, we admit that there could be
significant programmer effort involved in annotating the whole program. On the other hand,~\cite{nondex}
state "we found that manually inspecting these failures was
rather challenging, and we leave it as future work to automate
debugging test failures due to ADINS code". \TheDeterminismChecker reports source locations
which makes it easier for the programmer to fix issues and is orthogonal to the approach taken by NONDEX which
report failing tests.

@@

\section{deflaker case study}
\begin{table}[]
    \begin{tabular}{|l|l|}
        \hline
        Benchmark, issue & found by det checker? \\ \hline
        achilles, [2] issue \#309 &  Race condition\\ \hline
        achilles, [3] issue \#310 & Same fix for \#309 and \#310 \\ \hline
        checkstyle [4] issue \#4664& det checker not found \\ \hline
        jackrabbit-oak [6] \#6512&  Unresolved on issue tracker \\ \hline
        jackrabbit-oak [7] issue \#6524&  Dead link, didn't find 6524 in issue tracker \\ \hline
        nutz [10] issue \#1283 &  Removed tests (Printing (HTTP's) Response.getstatus())\\ \hline
        togglz [19] issue \#223 & (use NONDEX for debugging) \\ \hline
        togglz [20] issue \#240 &  \\ \hline
    \end{tabular}
\end{table}

>>>>

\todo{check if the bugs discovered by \theDeterminismChecker have test coverage}
\todo{We investigated the tests covering code that \theDeterminismChecker reported as nondeterministic.
If code is covered by tests, NonDex couldn't discover it. Else, its a limitation of NonDex.}
This indicates that the additional programmer effort required by \TheDeterminismChecker is justified
as our tool discovers non-trivial issues and doesn't depend on code coverage or other runtime heuristics.

@@

All the flakiness that NonDex found was due to 7 methods in 3 classes:
\<Class.getDeclaredFields>, \<Class.getDeclaredMethods>, \<Class.getFields>, \<DateFormatSymbols.getZoneStrings>,
\<HashMap.entrySet>, \<HashMap.keySet>, and \<HashMap.values>.  
We provide samples of the annotated source code form these benchmarks and the errors as reported by \TheDeterminismChecker 
in Figures~\ref{code-reflectasm} and~\ref{code-actiongenerator}.
\begin{figure}
    \begin{verbatim}
    static public FieldAccess get (Class type) {
        @Det ArrayList<@Det Field> fields = new ArrayList<Field>();
        ...
        while (nextClass != Object.class) {
            @Det Field @Det[] declaredFields = nextClass.getDeclaredFields();
            ...
        }
    }
    [ERROR] FieldAccess.java:[102,87] [assignment.type.incompatible] 
    incompatible types in assignment.
    found   : @Det Field @OrderNonDet []
    required: @Det Field @Det []
    \end{verbatim}
    \caption{Annotated source code from reflectasm.}
    \label{code-reflectasm}
\end{figure}
\begin{figure}
    \begin{verbatim}
    protected SimpleDataEvent createNextEvent() {
      for (@Det Entry<@Det String, @Det FieldType> entry : fields.entrySet()) {
          ...
      }
      ...
    }
    [ERROR] DataDictionaryEventSource.java:[73,78] [enhancedfor.type.incompatible] 
    incompatible types in enhanced for loop.
    found   : @NonDet Entry<@Det String, @Det FieldType>
    required: @Det Entry<@Det String, @Det FieldType>
    \end{verbatim}
    \caption{Annotated source code from ActionGenerator.}
    \label{code-actiongenerator}
\end{figure}
As can be seen from these figures, annotating local arrays and collections
has been sufficient to discover the sources of nondeterminism in most of the projects.
The rest of the projects required annotating the return types of methods.

For the benchmark \<commons-lang>, \cite{nondex} report multiple failing tests (14 flaky tests) due to calls
to \<Class.getDeclaredFields>. Upon inspecting the source code, we find that all of invocations to
\<Class.getDeclaredFields> are made from 5 locations. This means that annotating 5 lines of source code
would have been sufficient to identify the nondeterminism. Having said that, we admit that there could be
significant programmer effort involved in annotating the whole program. On the other hand,~\cite{nondex}
state "we found that manually inspecting these failures was
rather challenging, and we leave it as future work to automate
debugging test failures due to ADINS code". \TheDeterminismChecker reports source locations
which makes it easier for the programmer to fix issues and is orthogonal to the approach taken by NonDex which
report failing tests.

@@

\section{deflaker case study}
\begin{table}[]
    \begin{tabular}{|l|l|}
        \hline
        Benchmark, issue & found by det checker? \\ \hline
        achilles, [2] issue \#309 &  Race condition\\ \hline
        achilles, [3] issue \#310 & Same fix for \#309 and \#310 \\ \hline
        checkstyle [4] issue \#4664& det checker not found \\ \hline
        jackrabbit-oak [6] \#6512&  Unresolved on issue tracker \\ \hline
        jackrabbit-oak [7] issue \#6524&  Dead link, didn't find 6524 in issue tracker \\ \hline
        nutz [10] issue \#1283 &  Removed tests (Printing (HTTP's) Response.getstatus())\\ \hline
        togglz [19] issue \#223 & (use NonDex for debugging) \\ \hline
        togglz [20] issue \#240 &  \\ \hline
    \end{tabular}
\end{table}

### 2019-04-01 related.tex
Fix capitalization of NonDex

% https://github.com/TestingResearchIllinois/NonDex/wiki/Supported-APIs
NonDex uses a hand-crafted list of 47 methods (25 unique method names) in
in 13 classes.
This is analogous to our annotations on the JDK\@.
\todo{Say how much code our annotations cover?}
\todo{Check whether our JDK is missing anything they support.}
\todo{Did the NonDex authors contact the maintainers of the tools in which
  they found nondeterminism?}

>>>>

% https://github.com/TestingResearchIllinois/NonDex/wiki/Supported-APIs
NonDex uses a hand-crafted list of 47 methods (25 unique method names) in
in 13 classes.
This is analogous to our annotations on the JDK\@.
\todo{Say how much code our annotations cover?}
\todo{Check whether our JDK is missing anything they support.}
%\todo{Did the NonDex authors contact the maintainers of the tools in which
%  they found nondeterminism?}
\todo{The NonDex authors didn't contact the maintainers of tools.
From the paper: "NonDex detected 57 previously unknown
flaky tests in open-source projects and three flaky tests
that have been already fixed by the open-source software
developers."}
